An example exploratory analysis script

### 2024-04-06 R/analysis-code/exploratory-analysis.qmd
part 4 update

First, I am curious that if physical activity could be correlated to the choice of transportation since I have heard a theory that if someone take much effort on commute transportation, they tend to spend less time on exercise.

```{r}
# Fit multinomial logistic regression model
#library(nnet)
#multinom_model <- multinom(FAF ~ MTRANS, data = mydata)

# Summary of the model
#summary(multinom_model)
```
I will try to figure it out later how I could make it fitting.

@@

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors)) {
  subsets <- combn(predictors, i)
  for (j in 1:ncol(subsets)) {
    predictors_subset <- subsets[, j]

    # Fit model using subset of predictors
    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
    model <- lm(formula, data = trainData)

    # Make predictions on test set
    predictions <- predict(model, newdata = testData)

    # Calculate RMSE
    RMSE1 <- sqrt(mean((testData$BMI - predictions)^2))

    # Update minRMSE and bestSubset if RMSE is lower
    if (RMSE1 < minRMSE1) {
      minRMSE1 <- RMSE1
      bestSubset1 <- predictors_subset
    }
  }
}

@@

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors_updated)) {
  subsets <- combn(predictors_updated, i)
  for (j in 1:ncol(subsets)) {
    predictors_subset <- subsets[, j]

    # Fit model using subset of predictors
    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
    model <- lm(formula, data = trainData)

    # Make predictions on test set
    predictions <- predict(model, newdata = testData)

    # Calculate RMSE
    RMSE2 <- sqrt(mean((testData$BMI - predictions)^2))

    # Update minRMSE and bestSubset if RMSE is lower
    if (RMSE2 < minRMSE2) {
      minRMSE2 <- RMSE2
      bestSubset2 <- predictors_subset
    }
  }
}

@@

# Fit LASSO regression model using cross-validation
lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression

# Extract optimal lambda value
best_lambda <- lasso_model$lambda.min

# Extract coefficients of predictors for the optimal lambda
coef <- coef(lasso_model, s = best_lambda)

# Identify non-zero coefficients (selected predictors)
selected_predictors1 <- rownames(coef)[coef != 0]

# Print selected predictors
print(paste("Selected predictors:", paste(selected_predictors1, collapse = ", ")))
```

```{r}
# Fit LASSO regression model using cross-validation with interaction terms
x <- model.matrix(as.formula(paste(response, "~", paste(predictors_updated, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
y <- mydata[, response]
lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression

# Extract optimal lambda value
best_lambda <- lasso_model$lambda.min

# Extract coefficients of predictors for the optimal lambda
coef <- as.matrix(coef(lasso_model, s = "lambda.min"))

# Identify non-zero coefficients (selected predictors)
selected_predictors <- rownames(coef)[coef != 0]

# Print selected predictors
print(paste("Selected predictors (including interactions):", paste(selected_predictors, collapse = ", ")))
```

@@

# Define training control with repeated k-fold cross-validation
train_control <- trainControl(method = "repeatedcv", 
                              number = 10,       # Number of folds
                              repeats = 3)       # Number of repetitions

# Define the model formula
formula <- as.formula(paste(response, "~", paste(c(predictors, interaction_terms), collapse = " + ")))

# Train the model using cross-validation
model <- train(formula,
               data = mydata,
               method = "lm",                    # Use linear regression
               trControl = train_control,       # Use defined training control
               preProcess = c("center", "scale"),  # Preprocess data
               metric = "RMSE")                    # Use RMSE as the metric

@@

>>>>

First, I am curious that if physical activity could be correlated to the choice of transportation since I have heard a theory that if someone take much effort on commute transportation, they tend to spend less time on exercise.
First I want to use number to represent the type of transportation. Walking requires most energy consumption, then it goes with public transportation and automobile is the least.

```{r}
# Define a function to map transportation modes to numeric values
map_transport <- function(transport_mode) {
  if (transport_mode == "Walking") {
    return(4)
  } else if (transport_mode == "Motorbike") {
    return(2)
  } else if (transport_mode == "Public_Transportation") {
    return(3)
  } else if (transport_mode == "Automobile") {
    return(1)
  } else {
    return(0)  # Return 0 for other cases
  }
}

# Apply the function to create the new variable
mydata$MTRANSn <- sapply(mydata$MTRANS, map_transport)

# Calculate correlation coefficient
correlation <- cor(mydata$FAF, mydata$MTRANSn)

# Print correlation coefficient
print(correlation)
```
Since the correlation coefficient is close to zero, there is almost no linear association between them. So this result doesn't really support the theory that I have heard from social media.

@@

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors)) {
  subsets1 <- combn(predictors, i)
  for (j in 1:ncol(subsets1)) {
    predictors_subset1 <- subsets1[, j]

    # Fit model using subset of predictors
    formula_lm1 <- paste("BMI ~", paste(predictors_subset1, collapse = " + "))
    model_lm1 <- lm(formula_lm1, data = trainData)

    # Make predictions on test set
    predictions_lm1 <- predict(model_lm1, newdata = testData)

    # Calculate RMSE
    RMSE1 <- sqrt(mean((testData$BMI - predictions_lm1)^2))

    # Update minRMSE and bestSubset if RMSE is lower
    if (RMSE1 < minRMSE1) {
      minRMSE1 <- RMSE1
      bestSubset1 <- predictors_subset1
    }
  }
}

@@

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors_updated)) {
  subsets2 <- combn(predictors_updated, i)
  for (j in 1:ncol(subsets2)) {
    predictors_subset2 <- subsets2[, j]

    # Fit model using subset of predictors
    formula_lm2 <- paste("BMI ~", paste(predictors_subset2, collapse = " + "))
    model_lm2 <- lm(formula_lm2, data = trainData)

    # Make predictions on test set
    predictions_lm2 <- predict(model_lm2, newdata = testData)

    # Calculate RMSE
    RMSE2 <- sqrt(mean((testData$BMI - predictions_lm2)^2))

    # Update minRMSE and bestSubset if RMSE is lower
    if (RMSE2 < minRMSE2) {
      minRMSE2 <- RMSE2
      bestSubset2 <- predictors_subset2
    }
  }
}

@@

# Define number of folds for cross-validation
nfolds <- 10

# Perform cross-validation
cv_model <- cv.glmnet(x, y, alpha = 1, nfolds = nfolds)  # Use alpha = 1 for LASSO regression

# Extract selected predictors
selected_predictors <- rownames(coef(cv_model, s = "lambda.min"))[-1]  # Exclude intercept column

# Print selected predictors
print(paste("Selected predictors:", paste(selected_predictors, collapse = ", ")))
```

@@

# Define training control with repeated k-fold cross-validation
train_control1 <- trainControl(method = "repeatedcv", 
                              number = 10,       # Number of folds
                              repeats = 3)       # Number of repetitions

# Define the model formula
formula_CV <- as.formula(paste(response, "~", paste(c(predictors, interaction_terms), collapse = " + ")))

# Train the model using cross-validation
model <- train(formula_CV,
               data = mydata,
               method = "lm",                    # Use linear regression
               trControl = train_control,       # Use defined training control
               preProcess = c("center", "scale"),  # Preprocess data
               metric = "RMSE")                    # Use RMSE as the metric

@@

I will also try to use obesity level which is categorical variable so I can use DA models to fit the data and might compare the result with linear models. There are totally 7 categories classified by the original authors of the dataset.

```{r}
library(MASS)
library(caret)

# Define the outcome variable
outcome <- "Obesity"

# Define predictor variables
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")

# Create a training/testing split
set.seed(123)  # Set seed for reproducibility
trainIndex3 <- createDataPartition(mydata$Obesity, p = 0.8, list = FALSE)
trainData3 <- mydata[trainIndex, ]
testData3 <- mydata[-trainIndex, ]

# Initialize variables to store results
best_subset_DA <- NULL
best_accuracy_DA <- 0

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors)) {
  subsets3 <- combn(predictors, i)
  for (j in 1:ncol(subsets3)) {
    predictors_subset_DA <- subsets3[, j]

    # Fit the DA model using the subset of predictors
    formula_DA <- as.formula(paste(outcome, "~", paste(predictors_subset_DA, collapse = " + ")))
    model_DA <- lda(formula_DA, data = trainData3)

    # Make predictions on the testing set
    predicted_DA <- predict(model_DA, newdata = testData3)$class

    # Calculate accuracy
    accuracy_DA <- mean(predicted_DA == testData3$Obesity)

    # Update best subset and accuracy if current accuracy is higher
    if (accuracy_DA > best_accuracy_DA) {
      best_subset_DA <- predictors_subset_DA
      best_accuracy_DA <- accuracy_DA
    }
  }
}

# Print the best subset of predictors and accuracy
print(paste("Best subset of predictors:", paste(best_subset_DA, collapse = ", ")))
print(paste("Best accuracy:", best_accuracy_DA))
```

Adding 2 interactions to the DA model.

```{r}
# Define the outcome variable
outcome <- "Obesity"

# Create a training/testing split
set.seed(123)  # Set seed for reproducibility
trainIndex4 <- createDataPartition(mydata$Obesity, p = 0.8, list = FALSE)
trainData4 <- mydata[trainIndex, ]
testData4 <- mydata[-trainIndex, ]

# Create interaction terms in training and testing data
trainData4$Gender_Age <- trainData4$Gender * trainData4$Age
trainData4$FAF_Water <- trainData4$FAF * trainData4$Water

testData4$Gender_Age <- testData4$Gender * testData4$Age
testData4$FAF_Water <- testData4$FAF * testData4$Water

# Initialize variables to store results
best_subset_DA_U <- NULL
best_accuracy_DA_U <- 0

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors_updated)) {
  subsets4 <- combn(predictors_updated, i)
  for (j in 1:ncol(subsets4)) {
    predictors_subset_DA_U <- subsets4[, j]

    # Fit the DA model using the subset of predictors
    formula_DA_U <- as.formula(paste(outcome, "~", paste(predictors_subset_DA_U, collapse = " + ")))
    model_DA_U <- lda(formula_DA_U, data = trainData4)

    # Make predictions on the testing set
    predicted_DA_U <- predict(model_DA_U, newdata = testData4)$class

    # Calculate accuracy
    accuracy_DA_U <- mean(predicted_DA_U == testData4$Obesity)

    # Update best subset and accuracy if current accuracy is higher
    if (accuracy_DA_U > best_accuracy_DA_U) {
      best_subset_DA_U <- predictors_subset_DA_U
      best_accuracy_DA_U <- accuracy_DA_U
    }
  }
}

# Print the best subset of predictors and accuracy
print(paste("Best subset of predictors:", paste(best_subset_DA_U, collapse = ", ")))
print(paste("Best accuracy:", best_accuracy_DA_U))
```
After adding 2 interactions, the accuracy increased a little bit.

Using cross validation to evaluate the model.

```{r}
# Define predictors and response variable
mydata$Obesity <- factor(mydata$Obesity)
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
interaction_terms <- c("Gender_Age", "FAF_Water")
outcome <- "Obesity"

# Define the model formula
formula_CV_DA <- as.formula(paste(outcome, "~", paste(c(predictors, interaction_terms), collapse = " + ")))

# Define training control with repeated k-fold cross-validation
train_control2 <- trainControl(method = "repeatedcv", 
                               number = 10,            # Number of folds
                               repeats = 3,            # Number of repetitions
                               classProbs = TRUE,      # Compute class probabilities
                               summaryFunction = twoClassSummary)  # Use two-class summary

# Fit the DA model using cross-validation
lda_model <- train(formula = formula_CV_DA,
                   data = mydata,
                   method = "lda",             # Use LDA method
                   trControl = train_control2,  # Use defined training control
                   metric = "Accuracy",        # Use accuracy as the metric
                   preProcess = c("center", "scale")) # Preprocess data

print(lda_model)
```
